---
title: "Recreation PLOS Paper"
author: "Ulf Toelch"
date: "6 4 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(knitr)
library(kableExtra)
library(stringr)
# download the publication data from Figshare
url1 <- "https://ndownloader.figshare.com/files/9422038"
p1f <- tempfile()
download.file(url1, p1f, mode = "wb")
# read in data. Note that we need to define the NA
my_data <- read_excel(path = p1f, sheet = 1, na = "NA")
# the data set contains mortaility of patients after burns depending on
# whether and where they had inhalation burns
# first create a new variable that is a
# factor (0=normal, 1=subjective, 2=upper, and 3=lower)
my_data <-
  my_data %>%
  mutate(INHdiv_fac = factor(INHdiv, labels = c("normal", "subjective", "upper", "lower")))
```

## Recreation of some results
We will use a published data set here to recreate a table and create an advanced visualisation.
Does inhalation injury predict mortality in burns patients or require redefinition? see  https://doi.org/10.1371/journal.pone.0185195 for publication
the data set contains mortaility of patients after burns depending on whether and where they had inhalation burns

```{r table, message=F}
names_rows<-c("Mean age (years)","Sex (male:female)","Mean % TBSA burned")
names_columns<-c(paste("Total (n=",nrow(my_data),")",sep=""),paste("Survivors(n=",sum(my_data$mortaltiy==0),")"),paste("Non-survivors (n=",sum(my_data$mortaltiy==1),")"))

table_data<-
my_data %>% 
  summarise(mean_age=mean(age),sd_age=sd(age),
            sex_male=n()-sum(sex),sex_female=sum(sex),
            mean_TBSA=mean(TBSA),sd_TBSA=sd(TBSA)) %>%
  mutate(across(everything(),round,2)) %>% 
  unite(col="n_1",contains("age"),sep=" \u00b1 ") %>% 
  unite(col="n_2",contains("sex"),sep=" : ") %>% 
  unite(col="n_3",contains("TBSA"),sep=" \u00b1 ") %>% 
  pivot_longer(everything()) %>% 
  mutate(name=names_rows)
table_data %>% 
  kbl(col.names = c("",names_columns[1]),caption = "First Table") %>% 
  kable_paper("hover", full_width = F)

```
Now we try to add the next two columns.

```{r table2, message=F,echo=FALSE}
my_data<-
  my_data %>% 
  mutate(mortality=factor(mortaltiy,levels=c("Non-Survivors","Survivors")))

#next two columns
table_data2<-
  my_data %>% 
  group_by(mortaltiy) %>% 
  summarise(mean_age=mean(age),sd_age=sd(age),
            sex_male=n()-sum(sex),sex_female=sum(sex),
            mean_TBSA=mean(TBSA),sd_TBSA=sd(TBSA)) %>% 
  mutate(across(everything(),round,2)) %>% 
  unite(col="n_1",contains("age"),sep=" \u00b1 ") %>% 
  unite(col="n_2",contains("sex"),sep=" : ") %>% 
  unite(col="n_3",contains("TBSA"),sep=" \u00b1 ") %>% 
  pivot_longer(cols = contains("n")) %>% 
  pivot_wider(names_from=mortaltiy,values_from = value) %>% 
  mutate(name=names_rows)
table_data<-left_join(table_data,table_data2)

#use kable to create Table
#use proper colnames
table_data %>% 
  kbl(col.names = c("",names_columns[1:3]),caption="Second Table") %>% 
  kable_paper("hover", full_width = F)

```

The plots are not very appealing, we can do better....

```{r,message=FALSE,results='asis',echo=FALSE}
####################################
# create a data frame for a barplot
# group by the new inhalation severity variable and the PFdivide
# ("We also divided the patients into four groups depending on the PF ratio (>300, 200-300, 100-200, and <100)")
# you will need the mean of mortality (look out for spelling mistake) for each level of inhalation injury (INHdiv)
plot_data <-
  my_data %>%
  group_by(INHdiv_fac, Pfdivide) %>%
  summarise(
    mean_mort = mean(mortaltiy),
  )

# create a barplot from this with mean mortality on y axis and Pfdivide on x axis
# create a facet for each severity level
# add error bars
# make nice axis labels
p.1 <- ggplot(aes(y = mean_mort, x = Pfdivide), data = plot_data) +
  geom_bar(stat = "identity") +
  facet_wrap(~INHdiv_fac) +
  theme_classic() +
  ylab("Mean mortality") +
  xlab(expression("PaO"[2] * "/FiO"[2] * " (PF) ratio"))
print(p.1)
```

Now for the survival plots

```{r,results='asis',echo=FALSE,message=FALSE}
names(my_data)[8] <- c("PFratio")
hist_data <-
  my_data %>%
  # first add new variable breaks  that divides PFRatio in steps of 10
  # use the findInterval() function to assign the interval to each data point
  mutate(breaks = findInterval(PFratio, seq(20, 935, 10))) %>%
  # then group by dead/alive and the breaks
  group_by(mortaltiy, breaks) %>%
  # count the total number in the groups
  summarise(n = n()) %>%
  # if patients are dead, we want them to show on top with histogram on top so you need to
  # calculate in this case the percentage as 1-percentage
  mutate(
    pct = ifelse(mortaltiy == 0, n / sum(n), 1 - n / sum(n)),
    breaks = seq(20, 935, 10)[breaks]
  )
######
# now we create a the plot
#####
ggplot() + # this just sets an empty frame to build upon
  # first add a histopgram with geom_segment use the help of geom_segment
  geom_segment(
    data = hist_data, size = 2, show.legend = FALSE,
    aes(x = breaks, xend = breaks, y = mortaltiy, yend = pct, colour = factor(mortaltiy))
  ) +
  # then predict a logistic regression via stat_smooth and the glm method (we will cover the details in the next session)
  stat_smooth(data = my_data, aes(y = mortaltiy, x = PFratio), method = "glm", method.args = list(family = "binomial")) +
  # some cosmetics
  scale_y_continuous(limits = c(-0.02, 1.02)) +
  scale_x_continuous(limits = c(10, 950)) +
  theme_bw(base_size = 12) +
  ylab("Patient Alive=0/Dead=1") +
  xlab(expression("PaO"[2] * "/FiO"[2] * " (PF) ratio"))
######
# do the same thing for the four inhale burns groups separately
#####
hist_data <-
  my_data %>%
  # first add new variable that codes breaks
  mutate(breaks = findInterval(PFratio, seq(20, 935, 10))) %>%
  # then group by dead/alive and the breaks
  group_by(INHdiv_fac, mortaltiy, breaks) %>%
  # count
  summarise(n = n()) %>%
  # if patients are dead, we want them to show on top with histogram on top so you need to
  # calculate in this case the percentage as 1-percentage
  mutate(pct = ifelse(mortaltiy == 0, n / sum(n), 1 - n / sum(n)), breaks = seq(20, 935, 10)[breaks])
######
# again create the plot
#####
ggplot() + # this just sets an empty frame to build upon
  # set the four panels
  facet_wrap(~INHdiv_fac) +
  # first add a histopgram with geom_segment use the help of geom_segment
  geom_segment(
    data = hist_data, size = 2, show.legend = FALSE,
    aes(x = breaks, xend = breaks, y = mortaltiy, yend = pct, colour = factor(mortaltiy))
  ) +
  # then predict a logistic regression via stat_smooth and the glm method
  # (we will cover the details in the next session)
  stat_smooth(data = my_data, aes(y = mortaltiy, x = PFratio), method = "glm", method.args = list(family = "binomial")) +
  # some cosmetics
  scale_y_continuous(limits = c(-0.02, 1.02)) +
  scale_x_continuous(limits = c(10, 950)) +
  theme_bw(base_size = 12) +
  ylab("Patient Alive=0/Dead=1") +
  xlab(expression("PaO"[2] * "/FiO"[2] * " (PF) ratio"))

```

